---
title: "Project proposal"
author: "Community Concepts 2"
output: html_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
```

## 1. Introduction

**Community Concepts, INC. (CCI)** is an agency serving three counties in the state of Maine (Androscoggin, Franklin, and Oxford.) The agency helps provide stability and self-sufficiency through a variety of programs directed towards children, families, and households, such as early childhood education, housing, financial assistance, and energy assistance. 

In this work, we will be analyzing the efficiency of CCI's Child Services programs. These programs include **'Headstart Children Program'** and **'Early Headstart Children Program'**. The analysis will evaluate the total amount of USD invested in each program type by CCI, based on 2024 Town Campaign data, and attempt to find correlation between investment in each program and the efficiency (proximity to target set) of each program. Evaluation of targets will be collected through data on CCI's fiscal year 2024 Key Performance Indicators of Total Agency. 

Through finding correlations between the investment made in each program and that program's efficiency, we will have a better understanding of the ways in which investment can cause efficiency, or whether Community Concept's investments are not necessarily directly causing efficiency. This project aims to understand the inner workings of Community Concept's investment strategies, and can be a small glimpse of insight as to where CCI should continue or adjust investments in the future. 


### Research Question:

**How does funding and investment affect Children Services programs offered by Community Concepts such as Headstart and Early Headstart correlate with program targets, efficiency, and overall trends in Androscoggin, Franklin, and Oxford Counties in Maine?**

or

**How does funding for Community Concepts' Head Start and Early Head Start programs correlate with program efficiency and goal achievement in Androscoggin, Franklin, and Oxford Counties, Maine?**


## 2. Data

Our research question will be answered through the analysis of the folling variables:

- `id`: assigned ID
- `headstart`: headstart enrollment count
- `early_headstart`: early headstart enrollment count
- `monthly_enroll`: Cumulative enrollment
- `wait_income_eligible`: wait list for income eligible
- `applications`: applications processed within 10 days
- `hs_wl_incomes`:  HS Wait List for Income Eligible 
- `hs_wl_over`: HS Wait List for Over Income
- `ehs_wl_incomes`:  Early HS Wait List for Income Eligible 
- `ehs_wl_over`: Early HS Wait List for Over Income


In order to gain access to these variables for proper use, we must clean and effectively merge the Fiscal Year 2024 Key Perfomance Indicators data set with the Fiscal Year 24 Town Campaign Data, which includes all town names and the amount of investments raised in each campaign. 

The Town Campaign Data also must me narrowed down to only towns found in Androscoggin, Franklin, and Oxford counties, allowing for analysis of the efficiency in just the towns in which CCI covers, and therefore would recieve investment through.

The KPI set must be pivoted, in order for 'program' to be a column and 'month' to appear as rows.

Once the data is merged and cleaned, analysis and data visualization will help understand how funding and investment in Children Services programs offered by CCI correlate with program targets, efficiency, and overall trends.


```{r town import}
library(tidyverse)



library(readxl)
FY24_Town_Campaign_Data_1_ <- read_excel("/cloud/project/data/FY24 Town Campaign Data (1).xlsx",

    col_types = c("text", "skip", "skip", 
        "skip", "skip", "skip", "skip", "skip", 
        "skip", "skip", "skip", "skip", "skip", 
        "skip", "skip", "skip", "skip", "skip", 
        "skip", "skip", "skip", "skip", "skip", 
        "skip", "skip", "skip", "skip", "skip", 
        "text", "text", "text", "text", "text", 
        "text", "text", "skip", "skip", "skip", 
        "skip", "skip", "skip", "skip", "skip", 
        "skip", "skip", "skip", "skip", "skip", 
        "skip", "skip", "skip", "text"))
View(FY24_Town_Campaign_Data_1_)


```

```{r town column cleaning}
Town_data <- FY24_Town_Campaign_Data_1_

colnames(Town_data) <- as.character(unlist(Town_data[1, ]))
Town_data <- Town_data[-1, ]
Town_data <- Town_data[-(2:7), ]
Town_data <- Town_data[, -2]
 
 
colnames(Town_data)[2] <- "Headstart_individuals"
colnames(Town_data)[3] <- "Headstart_investments"

colnames(Town_data)[4] <- "EHS_individuals"
colnames(Town_data)[5] <- "EHS_investments"

colnames(Town_data)[6] <- "Nutrition_individuals"
colnames(Town_data)[7] <- "Nutrition_investments"

colnames(Town_data)[8] <- "Total"

```

```{r town deleting zeros}


Town_data[, c("Headstart_individuals", "Headstart_investments", 
              "EHS_individuals", "EHS_investments", 
              "Nutrition_individuals", "Nutrition_investments")] <- 
  lapply(Town_data[, c("Headstart_individuals", "Headstart_investments", 
                       "EHS_individuals", "EHS_investments", 
                       "Nutrition_individuals", "Nutrition_investments")], 
         function(x) replace(x, is.na(x), 0))

Town_data[, c("Headstart_individuals", "Headstart_investments",
              "EHS_individuals", "EHS_investments",
              "Nutrition_individuals", "Nutrition_investments")] <-
  lapply(Town_data[, c("Headstart_individuals", "Headstart_investments",
                       "EHS_individuals", "EHS_investments",
                       "Nutrition_individuals", "Nutrition_investments")], as.numeric)

Town_data <- Town_data[rowSums(Town_data[, c("Headstart_individuals", "Headstart_investments",
                                              "EHS_individuals", "EHS_investments",
                                              "Nutrition_individuals", "Nutrition_investments")]) != 0, ]

Town_data <- Town_data[-c(1, nrow(Town_data)), ]

```

```{r town totals sheet}

headstart_individuals_total <- sum(Town_data$Headstart_individuals)
headstart_investments_total <- sum(Town_data$Headstart_investments)

ehs_individuals_total <- sum(Town_data$EHS_individuals)
ehs_investments_total <- sum(Town_data$EHS_investments)

nutrition_individuals_total <- sum(Town_data$Nutrition_individuals)
nutrition_investments_total <- sum(Town_data$Nutrition_investments)

town_data_total <- data.frame(
  Headstart_individuals = headstart_individuals_total,
  Headstart_investments = headstart_investments_total,
  EHS_individuals = ehs_individuals_total,
  EHS_investments = ehs_investments_total,
  Nutrition_individuals = nutrition_individuals_total,
  Nutrition_investments = nutrition_investments_total
)


library(tidyr)

town_data_total <- data.frame(
  Program = c("Headstart", "Headstart", "Early Headstart", "Early Headstart", "Nutrition", "Nutrition"),
  Type = c("Individuals", "Investments", "Individuals", "Investments", "Individuals", "Investments"),
  Total = c(
    sum(Town_data$Headstart_individuals),
    sum(Town_data$Headstart_investments),
    sum(Town_data$EHS_individuals),
    sum(Town_data$EHS_investments),
    sum(Town_data$Nutrition_individuals),
    sum(Town_data$Nutrition_investments)
  )
)

town_data_total_long <- pivot_longer(town_data_total, 
                                     cols = c("Total"), 
                                     names_to = "Metric", 
                                     values_to = "Value")

print(town_data_total_long)


```


```{r KPI Import}
library(readxl)
FY24_Sept_KPIs_Total_Agency_1_ <- read_excel("/cloud/project/data/FY24 Sept KPIs - Total Agency (1).xlsx", 
    sheet = "Childrens ", col_types = c("text", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "skip", "skip", "skip", 
        "skip", "skip", "skip", "skip", "skip", 
        "skip", "skip", "skip", "skip"))
FY24_Sept_KPIs_Total_Agency_1_ 


# <- t(FY24_Sept_KPIs_Total_Agency_1_)
# colnames(FY24_Sept_KPIs_Total_Agency_1_) <- as.character(FY24_Sept_KPIs_Total_Agency_1_[1,])
# FY24_Sept_KPIs_Total_Agency_1_ <- FY24_Sept_KPIs_Total_Agency_1_[-1, ]  # Remove the first row now that it's been set as column names
# 
# rownames(FY24_Sept_KPIs_Total_Agency_1_)[12] <- "Sept" #rename to september
# FY24_Sept_KPIs_Total_Agency_1_ <- FY24_Sept_KPIs_Total_Agency_1_[, -19] #Remove column 18 because all values are 0
# FY24_Sept_KPIs_Total_Agency_1_ <- head(FY24_Sept_KPIs_Total_Agency_1_, -1)
# View(FY24_Sept_KPIs_Total_Agency_1_)
# 
# library(readxl)
# FY24_Sept_KPIs_Total_Agency_1_ <- read_excel("FY24 Sept KPIs - Total Agency (1).xlsx", 
#     sheet = "Childrens ", col_types = c("text", 
#         "numeric", "numeric", "numeric", 
#         "numeric", "numeric", "numeric", 
#         "numeric", "numeric", "numeric", 
#         "numeric", "numeric", "numeric", 
#         "numeric", "skip", "skip", "skip", 
#         "skip", "skip", "skip", "skip", "skip", 
#         "skip", "skip", "skip", "skip"))
# View(FY24_Sept_KPIs_Total_Agency_1_)
```

```{r KPI colnames}
KPI_data <- FY24_Sept_KPIs_Total_Agency_1_  

colnames(KPI_data)[1] <- "Program"  

colnames(KPI_data)[13] <- "Sept"

colnames(KPI_data)[14] <- "Total"

View(KPI_data)  

```



## 3. Data analysis plan

To effectively assess performance, we compare actual results against predefined goals across various Key Performance Indicators (KPIs). First, we prepare the data by creating a new column that calculates the percentage of the goal achieved, allowing for a standardized comparison across all metrics. Next, we visualize the data using bar plots or line charts to illustrate the differences between actual performance and target values, making it easier to identify trends and gaps. Finally, we conduct a statistical analysis to determine the average percentage of goal achievement, highlighting areas of strong performance as well as those that fall short. This approach provides a clear understanding of overall success while pinpointing specific areas that require strategic improvements.

```{r}
# Goal Achievement for Head Start Children Enrolled (Goal is 266)
children_values <- df$`Value`

#Example visualization
kpi_data <- kpi_data %>%
 mutate(goal_achievement = children_values /266 * 100)

ggplot(data, aes(x=month, y=goal_achievement)) +
 geom_bar(stat="identity") +
  labs(title="Goal Achievement for Head Start Children Enrolled", x="Month", y="Percentage of Goal Achieved")`

```


## 4. Data Ethics Review

The data ethics review section will be introduced in a separate class and is not part of the original proposal deadline.

