---
title: "Project memo"
author: "Team name"
output: github_document
---

This document should contain a detailed account of the data clean up for your data and the design choices you are making for your plots. For instance you will want to document choices you've made that were intentional for your graphic, e.g. color you've chosen for the plot. Think of this document as a code script someone can follow to reproduce the data cleaning steps and graphics in your handout.

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(viridis)
```

## Data Clean Up Steps for Overall Data

### Step 1: _________

```{r}

```

### Step 2: ________

## Plots

### ggsave example for saving plots

```{r starwars ggsave example}
p1 <- starwars |>
  filter(mass < 1000, 
         species %in% c("Human", "Cerean", "Pau'an", "Droid", "Gungan")) |>
  ggplot() +
  geom_point(aes(x = mass, 
                 y = height, 
                 color = species)) +
  labs(x = "Weight (kg)", 
       y = "Height (m)",
       color = "Species",
       title = "Weight and Height of Select Starwars Species",
       caption = paste("This data comes from the starwars api: https://swapi.py43.com"))


ggsave("example-starwars.png", width = 4, height = 4)

ggsave("example-starwars-wide.png", width = 6, height = 4)
```



### Plot 1: __Meals Distributed by Month_______

#### Data cleanup steps specific to plot 1

These data cleaning sections are optional and depend on if you have some data cleaning steps specific to a particular plot

```{r Clean_Total_Children}


KPI_data["Nov", "total_children"]<- 339
KPI_data["Dec", "total_children"]<- 341
KPI_data["Jan", "total_children"]<- 346
KPI_data["Feb", "total_children"]<- 364
KPI_data["Mar", "total_children"]<- 449
KPI_data["Apr", "total_children"]<- 452
KPI_data["May", "total_children"]<- 525
KPI_data["June", "total_children"]<- 525
KPI_data["July", "total_children"]<- 525
KPI_data["Aug", "total_children"]<- 525
KPI_data["Sept", "total_children"]<- 525

  
```


#### Final Plot 1

```{r}

```


### Plot 2: _________

### Plot 3: ___________

Add more plot sections as needed. Each project should have at least 3 plots, but talk to me if you have fewer than 3.

### Plot 4: ___________ 

