---
title: "Project memo"
author: "Community Concepts 2"
output: github_document
---

This document should contain a detailed account of the data clean up for your data and the design choices you are making for your plots. For instance you will want to document choices you've made that were intentional for your graphic, e.g. color you've chosen for the plot. Think of this document as a code script someone can follow to reproduce the data cleaning steps and graphics in your handout.

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(viridis)
```

## Data Clean Up Steps for Overall Data

### Step 1: Town Column Cleaning


```{r town column cleaning}
Town_data <- FY24_Town_Campaign_Data_1_

colnames(Town_data) <- as.character(unlist(Town_data[1, ]))
Town_data <- Town_data[-1, ]
Town_data <- Town_data[-(2:7), ]
Town_data <- Town_data[, -2]
 
 
colnames(Town_data)[2] <- "Headstart_individuals"
colnames(Town_data)[3] <- "Headstart_investments"

colnames(Town_data)[4] <- "EHS_individuals"
colnames(Town_data)[5] <- "EHS_investments"

colnames(Town_data)[6] <- "Nutrition_individuals"
colnames(Town_data)[7] <- "Nutrition_investments"

colnames(Town_data)[8] <- "Total"

```


### Step 2: Delete zeros

```{r town deleting zeros}


Town_data[, c("Headstart_individuals", "Headstart_investments", 
              "EHS_individuals", "EHS_investments", 
              "Nutrition_individuals", "Nutrition_investments")] <- 
  lapply(Town_data[, c("Headstart_individuals", "Headstart_investments", 
                       "EHS_individuals", "EHS_investments", 
                       "Nutrition_individuals", "Nutrition_investments")], 
         function(x) replace(x, is.na(x), 0))

Town_data[, c("Headstart_individuals", "Headstart_investments",
              "EHS_individuals", "EHS_investments",
              "Nutrition_individuals", "Nutrition_investments")] <-
  lapply(Town_data[, c("Headstart_individuals", "Headstart_investments",
                       "EHS_individuals", "EHS_investments",
                       "Nutrition_individuals", "Nutrition_investments")], as.numeric)

Town_data <- Town_data[rowSums(Town_data[, c("Headstart_individuals", "Headstart_investments",
                                              "EHS_individuals", "EHS_investments",
                                              "Nutrition_individuals", "Nutrition_investments")]) != 0, ]

Town_data <- Town_data[-c(1, nrow(Town_data)), ]

```

### Step 3 Create town totals

```{r town totals sheet}

headstart_individuals_total <- sum(Town_data$Headstart_individuals)
headstart_investments_total <- sum(Town_data$Headstart_investments)

ehs_individuals_total <- sum(Town_data$EHS_individuals)
ehs_investments_total <- sum(Town_data$EHS_investments)

nutrition_individuals_total <- sum(Town_data$Nutrition_individuals)
nutrition_investments_total <- sum(Town_data$Nutrition_investments)

town_data_total <- data.frame(
  Headstart_individuals = headstart_individuals_total,
  Headstart_investments = headstart_investments_total,
  EHS_individuals = ehs_individuals_total,
  EHS_investments = ehs_investments_total,
  Nutrition_individuals = nutrition_individuals_total,
  Nutrition_investments = nutrition_investments_total
)


library(tidyr)

town_data_total <- data.frame(
  Program = c("Headstart", "Headstart", "Early Headstart", "Early Headstart", "Nutrition", "Nutrition"),
  Type = c("Individuals", "Investments", "Individuals", "Investments", "Individuals", "Investments"),
  Total = c(
    sum(Town_data$Headstart_individuals),
    sum(Town_data$Headstart_investments),
    sum(Town_data$EHS_individuals),
    sum(Town_data$EHS_investments),
    sum(Town_data$Nutrition_individuals),
    sum(Town_data$Nutrition_investments)
  )
)

town_data_total_long <- pivot_longer(town_data_total, 
                                     cols = c("Total"), 
                                     names_to = "Metric", 
                                     values_to = "Value")

print(town_data_total_long)


```

### Step 4: Create Columns for KPI dataset

```{r KPI colnames}
KPI_data <- FY24_Sept_KPIs_Total_Agency_1_  

colnames(KPI_data)[1] <- "Program"  

colnames(KPI_data)[13] <- "Sept"

colnames(KPI_data)[14] <- "Total"

View(KPI_data)  

```




## Plots

### ggsave example for saving plots

```{r starwars ggsave example}
p1 <- starwars |>
  filter(mass < 1000, 
         species %in% c("Human", "Cerean", "Pau'an", "Droid", "Gungan")) |>
  ggplot() +
  geom_point(aes(x = mass, 
                 y = height, 
                 color = species)) +
  labs(x = "Weight (kg)", 
       y = "Height (m)",
       color = "Species",
       title = "Weight and Height of Select Starwars Species",
       caption = paste("This data comes from the starwars api: https://swapi.py43.com"))


ggsave("example-starwars.png", width = 4, height = 4)

ggsave("example-starwars-wide.png", width = 6, height = 4)
```



### Plot 1 and 2: Town Totals: Investments and Invidiviuals


#### Final Plot 1 and 2


```{r town plots, investments and individuals}

library(tidyr)
library(ggplot2)
library(scales)

town_data_total_long$Program <- factor(town_data_total_long$Program, 
                                       levels = c("Headstart", "Early Headstart", "Nutrition"))

individuals_plot <- ggplot(town_data_total_long[town_data_total_long$Type == "Individuals", ], 
                           aes(x = Program, y = Value, fill = Program)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Individuals per Program", x = "Program", y = "Number of Individuals") +
  theme_minimal()

investments_plot <- ggplot(town_data_total_long[town_data_total_long$Type == "Investments", ], 
                           aes(x = Program, y = Value, fill = Program)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Investments per Program", x = "Program", y = "Investment Value") +
  scale_y_continuous(labels = label_comma()) +  
  theme_minimal()

individuals_plot
investments_plot


```


### Plot 3: KPI Attendance Goals Vs. Actual Attendance

```{r KPI enrollment plot}

KPI_data <- as_tibble(KPI_data)

KPI_enrollment <- KPI_data %>%
  slice(2, 7)

KPI_enrollment_long <- KPI_enrollment %>%
  pivot_longer(cols = -Program, names_to  = "Month", values_to = "Enrollment") %>%
  filter(!is.na(Enrollment)) %>%  
  mutate(Month = factor(Month,levels = c("Oct", "Nov", "Dec", "Jan", "Feb", "Mar", "Apr", "May", "June", "July", "Aug", "Sept")),
    Program = recode(Program,
      "HS Goal ≥ 90%   Attendance %"  = "Head Start",
      "EHS Goal ≥ 90% Attendance %"   = "Early Head Start"))
KPI_enrollment_long <- KPI_enrollment_long[-c(13, 26), ]

ggplot(KPI_enrollment_long,
       aes(x = Month, y = Enrollment,
           group = Program, color = Program)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  geom_hline(yintercept = 0.90,
             linetype   = "solid",
             color      = "red") +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    limits = c(0.5, 1.0)) +
  labs(
    title    = "Attendance Percentages vs 90% Goal",
    subtitle = paste0(
      "The percentage of attendance to Head Start and Early Head Start\n",
      "programs by month, compared to CCI's 90% attendance goal (2024)"),
    x = "Month",
    y = "Enrollment Rate, Percent") +
  theme_minimal()

```
### Plot 4: KPI Waitlist Plot

```{r KPI Plot}
KPIs_Total_Agency$month <- fct_relevel(KPIs_Total_Agency$month, "jan", "feb", "mar", "apr", "may", "june", "july", "aug", "sep", "oct", "nov", "dec")

# Reshape data for stacked bar plot
data_waitlist_melted <- melt(KPIs_Total_Agency, id.vars = "month")

# Stacked bar plot for Waitlist Income and Overload
ggplot(data_waitlist_melted, na.rm = TRUE, aes(x = factor(month, levels = c("jan", "feb", "mar", "apr", "may", "june", "july", "aug", "sep", "oct", "nov", "dec")), y = value, fill = variable)) +
geom_bar(stat = "identity") +
labs(title = "Waitlist Income and Overload", x = "Month", y = "Count") +
theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
scale_fill_manual(values = c("hs_wl_income" = "skyblue", "hs_wl_over" = "orange", "ehs_wl_income" = "lightgreen", "ehs_wl_over" = "red"),
    labels = c("HS Wait List for Income Eligible", "HS Wait List for Over Income", "Early HS Wait List for Income Eligible", "Early HS Wait List for Over Income"))# Custom legend labels
```

